---
title: "Bootstrapping"
author: "Chhiring Lama"
date: "2024-11-14"
output: github_document
---

```{r setup, include=FALSE, message = FALSE}
library(tidyverse)
library(modelr)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "85%", 
	fig.align = "center"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

Getting started with bootstrapping

Lets make up some data
```{r}
n_samp = 200

sim_df_constant <- tibble(x = rnorm(n_samp,1, 1), 
         error =rnorm(n_samp, 0, 1), 
         y = 2 + 3*x + error)
  

sim_df_nonconstant <- sim_df_constant |> 
  mutate(error = error*0.75*x, 
         y = 2 + 3*x + error)
```

Lets look at the dataset.
```{r}
sim_df_constant |> 
  ggplot(aes(x=x, y =y)) +
  geom_point() +
  stat_smooth(method = "lm")

sim_df_nonconstant |> 
  ggplot(aes(x=x, y =y)) +
  geom_point() +
  stat_smooth(method = "lm")
```

Look at regression results
```{r}
sim_df_constant |> 
  lm(y ~ x, data = _) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)

sim_df_nonconstant |> 
  lm(y ~ x, data = _) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```

Despite the very different error structures, standard errors for coefficient estimates are similar in both cases!

We’ll use the bootstrap to make inference for the data on the right. This is intended largely as an illustration for how to use the bootstrap in cases where the theoretical distribution is “unknown”, although for these data in particular weighted least squares could be more appropriate.

## Draw a bootstrap sample

```{r}
boot_sample <- function(df){
  boot_df = sample_frac(df, replace = TRUE) |> 
    arrange(x)
  return(boot_df)
}
```

Lets try running this
```{r}
sim_df_nonconstant |> 
  boot_sample() |> 
  ggplot(aes(x=x, y =y)) +
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm")
```
Each bootstrap gives us a slightly different fit. 

Can we do this as part of an analysis? Yes
```{r}
sim_df_nonconstant |> 
  boot_sample() |> 
  lm(y ~ x, data = _) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```

We can save the estimated values in each bootstrap sampling
```{r}
boot_straps <- tibble(
  strap_number = 1:1000
) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = sim_df_nonconstant)), 
    models = map(strap_sample, \(df) lm(y ~ x, data = df)), 
    results = map(models, broom::tidy)
  )

bootstrap_results <- boot_straps |> 
  select(strap_number, results) |> 
  unnest(results) |> 
  group_by(term) |> 
  summarise(
    boot_se = sd(estimate)
  )
```

We can try it for the sample where the assumptions are not violated (`sim_df_constant`). We get the same SE in both, meaning bootstrapping holds the theory of large numbers for both sample adhering to and violating the normality assumption. 

## Do this using modelr

```{r}
boot_straps <- sim_df_nonconstant |> 
  modelr::bootstrap(1000) |> 
  mutate(strap = map(strap, as_tibble), 
         models = map(strap, \(df) lm(y ~ x, data = df)), 
    results = map(models, broom::tidy)) |> 
  select(.id, results) |> 
  unnest(results)
```

## What do you want to report

```{r}
boot_straps |> 
  summarise(boot_se = sd(estimate), 
            boot_est = mean(estimate), 
            boot_ci_ll = quantile(estimate, 0.025), 
            boot_ci_ul = quantile(estimate, 0.975))
```

## Airbnb

```{r}
data("nyc_airbnb")

manhattan_airbnb <-  
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough == "Manhattan") |> 
  select(price, stars, borough, neighborhood, room_type) |> 
  drop_na()
```

plot the data
```{r}
manhattan_airbnb |> 
  ggplot(aes(x = stars, y = price)) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)
```

Fit a regression
```{r}
manhattan_airbnb |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```

Bootstrap for better inference?

```{r}
bootstrap_results <- manhattan_airbnb |> 
  modelr::bootstrap(1000) |> 
  mutate(
    strap = map(strap, as_tibble), 
    models = map(strap, \(df) lm(price ~ stars + room_type, data = df)), 
    results = map(models, broom::tidy)
  ) |> 
  select(.id, results) |> 
  unnest(results)
```

```{r}
bootstrap_results |> 
  filter(term == "stars") |> 
  ggplot(aes(x = estimate)) +
  geom_density()
```

```{r}
bootstrap_results |> 
  summarise(boot_se = sd(estimate), 
            boot_est = mean(estimate), 
            boot_ci_ll = quantile(estimate, 0.025), 
            boot_ci_ul = quantile(estimate, 0.975))
```











