---
title: "linear_models"
author: "Chhiring Lama"
date: "2024-11-07"
output: github_document
---

```{r setup, include=FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "85%", 
	fig.align = "center"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

Import the dataset and do some data cleaning:
```{r}
data("nyc_airbnb")

nyc_airbnb <-  
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type) |> 
  mutate(borough = fct_infreq(borough), 
         room_type = fct_infreq(room_type))
    
```

Fit linear regression to predict price using stars and borough

Relatively simple:
```{r}
fit <- lm(price ~ stars + borough, data = nyc_airbnb)

summary(fit)

names(summary(fit))

coef(fit)

fit |> 
  broom::tidy() |> 
  select(term, estimate, `p.value`)

fit |> 
  broom::glance()
```

Try a more complex model
```{r}
fit <- lm(price ~ stars + borough, data = nyc_airbnb)

fit |> 
  broom::tidy() |> 
  select(term, estimate, `p.value`) |> 
  mutate(term = str_replace(term, "borough", "Borough:")) |> 
  knitr::kable(digits = 3)
```

## Some diagnostics

(backtrack to some EDA)

```{r}
nyc_airbnb |> 
  ggplot(aes(y = price, x = stars)) +
  geom_point() +
  stat_smooth(method = "lm")
```


Most diagnostics use residuals
```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = resid)) +
  geom_histogram()
```

```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin() +
  ylim(-200, 500)
```

Residuals against stars
```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```

Residuals against fitted values
```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point()
```

## Hypothesis Testing
For single coefficient, at the table
```{r}
fit |> 
  broom::tidy()
```

When adding more than one coeff, use anova(), 
```{r}
fit_null <- lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt <- lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```

Look at both
```{r}
fit_null |> 
  broom::tidy()
fit_alt |> 
  broom::tidy()

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

## Do effects differ across boroughs

First, use a lot of interactions
```{r}
nyc_airbnb |> 
  lm(price ~ stars + borough + room_type*borough, data = _) |> 
  broom::tidy()
```
Strong interaction between borough and room type. But, we dont the situation in manhattan and roomtypes. 

Could fit separate models. 
```{r}
nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```
Within, Manhattan, strong association with roomtypes for price. 

```{r}
nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```
Similar observation here in Brooklyn too. Can check for the Bronx and Queens too. 

Get fancy and use list columns
```{r}
nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)), 
    results = map(model, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term, 
    values_from = estimate
  )
```

What about room type across manhattan neighborhoods?

```{r}
nyc_airbnb |> 
  filter(borough == "Manhattan", 
         neighborhood == "Chinatown") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Manhattan", 
         neighborhood == "Chelsea") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
  
```

Lets do this for all neighborhoods!
```{r}
manhattan_neighborhood_fit_df <- nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  nest(data = -(borough:neighborhood)) |> 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)), 
    results = map(model, broom::tidy)
  ) |> 
  select(borough, neighborhood, results) |> 
  unnest(results) 
```

Look at the effect of the room type
```{r}
manhattan_neighborhood_fit_df |> 
  filter(str_detect(term, "room_type")) |> 
  ggplot(aes(x = term, y = estimate)) +
  geom_boxplot()

nyc_airbnb |> 
  filter(neighborhood == "NoHo", 
         room_type == "Shared room")
```

## Binary outcome
```{r}
baltimore_df <-  
  read_csv("data/homicide-data.csv") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) |> 
  select(resolved, victim_age, victim_race, victim_sex)
```

```{r}
fit_logistic = 
  baltimore_df |> 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) 
```

```{r}
fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate)) |>
  select(term, log_OR = estimate, OR, p.value) |> 
  knitr::kable(digits = 3)
```

We can also compute fitted values; similarly to the estimates in the model summary, these are expressed as log odds and can be transformed to produce probabilities for each subject.
```{r}
baltimore_df |> 
  modelr::add_predictions(fit_logistic) |> 
  mutate(fitted_prob = boot::inv.logit(pred))
```









